\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-10-24 gwnnDraft02.tex
final changes
PRD updated submission

In this work, we explore the possibility of having multiple decoupled hidden sectors. Large numbers of hidden sectors can solve the hierarchy problem as in the Dvali Redi model~\cite{Dvali:2009ne}, in the more recently explored $N$naturalness~\cite{Arkani-Hamed:2016rle} framework, or in orbifold Higgs models~\cite{Craig:2014aea,Craig:2014roa}. They can also be motivated by dark matter considerations~\cite{Dienes:2011ja,Dienes:2011sa,Dienes:2016vei}. 
Motivated by solutions to the hierachy problem, we consider hidden sectors with the same particle content as the Standard Model that have all dimensionless couplings (defined at some high scale) equal to those of the Standard Model. The only parameter that varies across sectors is the dimension two Higgs mass squared parameter, $m_H^2$. This simple ansatz can lead to very rich phenomenology and interesting gravitational wave spectra, but we stress that it is only a starting point for exploring the space of theories with multiple hidden sectors.

@@

Gravitational waves are produced with contributions from different components of the SFOPT's evolution.  It is commonplace to parameterize the spectral energy density in gravitational waves by~\citep{PhysRevD.75.043507} 
\begin{equation}\label{eqn:SpectralEng}
\Omega_{\textrm{GW}} (f) \equiv \frac{1}{ \rho_{c}} \frac{d \rho_{\textrm{GW}}(f) }{d \textrm{\,log}(f)} , 
\end{equation}
where $\rho_{c} = 3H^2/(8 \pi G)$ is the critical energy density. The total gravitational wave signal is a linear combination of three leading contributions:
\begin{equation}
h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi} + h^2\Omega_{v} + h^2\Omega_{turb} .
\end{equation}
The three leading order contributions to the GW power spectrum are as follows:
\begin{itemize}
\item \textbf{Scalar field contributions $\Omega_{\phi}$}: Caused by collisions of the bubble walls, the solutions being completely dependent on the scalar field configuration, with efficiency factor $\kappa_{\phi} = 1 - \alpha_{\infty}/\alpha$~\citep{PhysRevD.45.4514, Huber_2008}. 
\item \textbf{Sound wave contributions $\Omega_{v}$}: Sound waves within the plasma after bubble collision will produce $\beta/H$ enhanced gravitational waves, with efficiency factor $\kappa_{v} \propto \alpha_{\infty}/\alpha$~\citep{PhysRevLett.112.041301}.
\item \textbf{Magnetohydrodynamical contributions $\Omega_{B}$}: Turbulence within the plasma, left over from the sound wave propagation, will produce gravitational waves with efficiency factor $\kappa_{
turb} \approx 0.1 \kappa_{v} $~\citep{PhysRevD.74.063521}.
\end{itemize}
The parameter $\alpha_{\infty}$ denotes the dividing line between the runaway regime  $(\alpha >\alpha_{\infty})$ and the non-runaway regime $(\alpha <\alpha_{\infty})$. Explicitly \cite{Breitbach:2018ddu, Caprini:2015zlo, Espinosa:2010hh}, 
\begin{equation}\label{eqn:critPTstrength}
\alpha_{\infty} = \frac{(T^{nuc}_h)^2}{\rho_R}\left[\sum_{bosons} n_i\frac{\Delta m^2_i}{24} + \sum_{fermions} n_i\frac{\Delta m_i^2}{48}\right],
\end{equation} 
for particles with $n_i$ degrees of freedom that obtain mass through the phase transition.

@@

Shifting to the other part of our framework, our results leads to the question ``how well can experiments probe non power law signals?" For frequency ranges away from the peak of the total, GW signals the quoted detection thresholds should hold: the signals fall off as a power law to a very good approximation. However, for areas around the peak frequency the answer is less clear; the PLI curves are built under the assumption of a power law. This points to the need for future work to better understand detection prospects for the next generation of GW detectors.

>>>>

In this work, we explore the possibility of having multiple decoupled hidden sectors. Large numbers of hidden sectors can solve the hierarchy problem as in the Dvali Redi model~\cite{Dvali:2009ne}, in the more recently explored $N$naturalness~\cite{Arkani-Hamed:2016rle} framework, or in orbifold Higgs models~\cite{Craig:2014aea,Craig:2014roa}. They can also be motivated by dark matter considerations~\cite{Chialva_2013,Dienes:2011ja,Dienes:2011sa,Dienes:2016vei}. 
Motivated by solutions to the hierachy problem, we consider hidden sectors with the same particle content as the Standard Model that have all dimensionless couplings (defined at some high scale) equal to those of the Standard Model. The only parameter that varies across sectors is the dimension two Higgs mass squared parameter, $m_H^2$. This simple ansatz can lead to very rich phenomenology and interesting gravitational wave spectra, but we stress that it is only a starting point for exploring the space of theories with multiple hidden sectors.

@@

Gravitational waves are produced with contributions from different components of the SFOPT's evolution.  It is commonplace to parameterize the spectral energy density in gravitational waves by~\citep{PhysRevD.75.043507} 
\begin{equation}\label{eqn:SpectralEng}
\Omega_{\textrm{GW}} (f) \equiv \frac{1}{ \rho_{c}} \frac{d \rho_{\textrm{GW}}(f) }{d \textrm{\,log}(f)} , 
\end{equation}
where $\rho_{c} = 3H^2/(8 \pi G)$ is the critical energy density. The total gravitational wave signal is a linear combination of three leading contributions:
\begin{equation}
h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi} + h^2\Omega_{v} + h^2\Omega_{turb} .
\end{equation}
Each components is scaled by its own unique efficiency factor, $\kappa$. The three leading order contributions to the GW power spectrum are as follows:
\begin{itemize}
\item \textbf{Scalar field contributions $\Omega_{\phi}$}: Caused by collisions of the bubble walls, the solutions being completely dependent on the scalar field configuration, with efficiency factor $\kappa_{\phi} = 1 - \alpha_{\infty}/\alpha$~\citep{PhysRevD.45.4514, Huber_2008}. 
\item \textbf{Sound wave contributions $\Omega_{v}$}: Sound waves within the plasma after bubble collision will produce $\beta/H$ enhanced gravitational waves, with efficiency factor $\kappa_{v} \propto \alpha_{\infty}/\alpha$~\citep{PhysRevLett.112.041301}.
\item \textbf{Magnetohydrodynamical contributions $\Omega_{B}$}: Turbulence within the plasma, left over from the sound wave propagation, will produce gravitational waves with efficiency factor $\kappa_{
turb} \approx 0.1 \kappa_{v} $~\citep{PhysRevD.74.063521}.
\end{itemize}
The parameter $\alpha_{\infty}$ denotes the dividing line between the runaway regime  $(\alpha >\alpha_{\infty})$ and the non-runaway regime $(\alpha <\alpha_{\infty})$. Explicitly \cite{Breitbach:2018ddu, Caprini:2015zlo, Espinosa:2010hh}, 
\begin{equation}\label{eqn:critPTstrength}
\alpha_{\infty} = \frac{(T^{nuc}_h)^2}{\rho_R}\left[\sum_{bosons} n_i\frac{\Delta m^2_i}{24} + \sum_{fermions} n_i\frac{\Delta m_i^2}{48}\right],
\end{equation} 
for particles with $n_i$ degrees of freedom that obtain mass through the phase transition.

@@

Shifting to the other part of our framework, our results leads to the question ``how well can experiments probe non power law signals?" For frequency ranges away from the peak of the total, GW signals the quoted detection thresholds should hold: the signals fall off as a power law to a very good approximation. However, for areas around the peak frequency the answer is less clear; the PLI curves are built under the assumption of a power law. Work has been done \cite{alanne2019fresh} in examining GW signals using peak amplitudes and peak frequencies as the defining observables: this is rooted in the assumption that GW signals have a model-independent spectral shape. However, our results indicate that this assumption of model-independence cannot hold for all cases: sectors with similar (but different) transition temperatures can create either peak broadening or multi-hump features that differ significantly from a standard power law shape.  This points to the need for future work to better understand where the power law approximation breaks down and how this affects detection prospects for the next generation of GW detectors.
